External Representation of Spatial and Spatio-Temporal Values

### 2004-05-03 Documents/STFormat.tex
Time representation revised

\section{Representation of Time}
There are several representations for an \emph{instant} value.
All Applications using instants must be able to read each of this.
A \emph{DateTime} as an algebra type consist only of its value list. 
If \emph{DateTime} is used as subtype, e.g. in an interval, the
type-name has to be included. The typename can be {\tt datetime} or
{\tt instant}. Summarized we have three possibilities for a \emph{DateTime}:
\begin{tt}
  \begin{small}
     \begin{itemize}
       \item (datetime $<$datetimevalue$>$) \\
       \item (instant $<$datetimevalue$>$)  \\
       \item $<$datetimevalue$>$
      \end{itemize}
  \end{small}
\end{tt}

The representations of {\tt $<$datetimevalue$>$} are described
in the next subsections.

\subsection{String Representation}
\label{StringRepOfTime}
A $<$datetimevalue$>$ can be represented as string in the following
format: \\
\begin{tt}
  \begin{small}
      year-month-day[-hour:minute[:second[.millisecond]]]
  \end{small}
\end{tt}\\
The squared brackets indicate optionally parts. All contained values
are integers which must be in the appropriate interval.

\subsection{Real Representation}
A $<$datetimevalue$>$ can just be represented as an real value. 
The conversion beween the formats is described in section 
\ref{TimeConversions}.

\subsection{Julian Representation}
A $<$datetimevalue$>$ in the Julian representation is:\\
\begin{tt}
  \begin{small}
      (day millisecond)
  \end{small}
\end{tt}\\
{\tt day} can be any integer and millisecond is an integer in 
the interval [0,86400000[. {\tt day} represents the difference
to a fixed nulldate and {\tt milliseconds} represents the 
time part of this day.

\subsection{Gregorian Representation}
In the Gregorian Representation a $<$datetimevalue$>$ is written 
down as follow:\\
\begin{tt}
  \begin{small}
      (day month year [ hour minute [ second [ millisecond]]])
  \end{small}
\end{tt}\\
Like in the string representation (see section \ref{StringRepOfTime}),
the squared brackets are not part of the representation but indicates
optionally information. The values of the omitted parts are assumed to be zero.
The triple (day, month, year) must be a valid date. The remaining
values must be in the appropriate interval.

\subsection{Conversions between Representations of Time}
\label{TimeConversions}
The conversion between {\tt real} value and Julian representation
is straightforward:\\
$ JulToReal : int \times int \rightarrow \R$ \\
$ JulToReal(day,millisecond) = day + \frac{millisecond}{86400000}$ \\
The used operators are working on real numbers.\\[1em]
The another direction is computed as follow. Note a possible lost of 
precision.\\
$ RealToJul : \R \rightarrow (int,int)$ \\
$ RealToJul(instant) = (day,milli) $ \\
where:\\
{\small
$ day = \left\{ \begin{array}{ll}
                    int(instant) & \mbox{if } instant > 0 
                                   \mbox{ or } int(instant)=instant\\
                    int(instant)-1 & otherwise
                \end{array} \right. $\\[1em]
$ milli = \left\{ \begin{array}{ll}
                         tmp & \mbox{if } instant > 0 \mbox{ or } int(instant)=instant\\
                         tmp + 86400000 & \mbox{otherwise}
                      \end{array}\right. $ \\[1em]
\hspace*{1cm}where : tmp = int ((instant-int(instant))*86400000+0.5)}\vspace{1cm}

>>>>

\section{Representation of Time}
They are two types of time. The first one is {\tt instant}, and the 
second one is {\tt duration}. If a time value is used internal (e.g.
in moving objects), the list will be:\\
  {\tt ( type $<$timevalue$>$ )}\\
where {\tt type} is a Symbol atom which can holds one of the values
{\tt\{ instant, duration\}}. For the reason of compatibility we
allow also the value {\tt datetime} as an alias for an instant type.

The representations of {\tt $<$timevalue$>$} are described
in the next subsections.

\subsection{String Representation}
\label{StringRepOfTime}
A $<$timevalue$>$ of type {\tt instant} can be represented as 
string in the following format: \\
\begin{tt}
  \begin{small}
      year-month-day[-hour:minute[:second[.millisecond]]]
  \end{small}
\end{tt}\\
The squared brackets indicate optionally parts. All contained values
are integers which must be in the appropriate interval.

\subsection{Real Representation}
A $<$timevalue$>$ can just be represented as an real value. 
The conversion beween the formats is described in section 
\ref{TimeConversions}. This format is accepted by {\tt instant} and
{\tt duration} types.

\subsection{Julian Representation}
A $<$timevalue$>$ in the Julian representation is defined for {\tt duration} 
types as follow:\\
\begin{tt}
  \begin{small}
      (day millisecond)
  \end{small}
\end{tt}\\
{\tt day} can be any integer and millisecond is an integer in 
the interval [0,86400000[.

\subsection{Gregorian Representation}
In the Gregorian Representation (for {\tt instant}s only)
a $<$timevalue$>$ is written 
down as:\\
\begin{tt}
  \begin{small}
      (day month year [ hour minute [ second [ millisecond]]])
  \end{small}
\end{tt}\\
Like in the string representation (see section \ref{StringRepOfTime}),
the squared brackets are not part of the representation but indicates
optionally information. The values of the omitted parts are assumed to be zero.
The triple (day, month, year) must be a valid date. The remaining
values must be in the appropriate interval.

\subsection{Conversions between Representations of Time}
\label{TimeConversions}
Both types of time can be represented by a tuple {\tt (day, milliseconds)}. 
The Meaning for a duration type is clear. For an instant type this tuple
is the difference to a fixed {\tt NULL\_DATE}. Because this {\tt NULL\_DATE}
is given in the Julian calender, we denote this representation as 
Julian Representation.
The conversion between {\tt real} value and Julian representation
is straightforward:\\
$ JulToReal : int \times int \rightarrow \R$ \\
$ JulToReal(day, millisecond) = day + \frac{millisecond}{86400000}$ \\
The used operators are working on real numbers.\\[1em]
The another direction is computed as follow. Note a possible lost of 
precision.\\
$ RealToJul : \R \rightarrow (int,int)$ \\
$ RealToJul(instant) = (day,milli) $ \\
where:\\
{\small
$ day = \left\{ \begin{array}{ll}
                    int(instant) & \mbox{if } instant > 0 
                                   \mbox{ or } int(instant)=instant\\
                    int(instant)-1 & otherwise
                \end{array} \right. $\\[1em]
$ milli = \left\{ \begin{array}{ll}
                         tmp & \mbox{if } instant > 0 \mbox{ or } int(instant)=instant\\
                         tmp + 86400000 & \mbox{otherwise}
                      \end{array}\right. $ \\[1em]
\hspace*{1cm}where : tmp = int ((instant-int(instant))*86400000+0.5)}\vspace{1cm}
